Toward a novel piezoelectric energy harvester for earcanal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes

### 2022-11-03 article_system.tex
Après correction anglais ETS

\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{graphicx}

@@

\begin{abstract}
Scavenging energy from the earcanal's dynamic motion during jaw movements may be a practical way to enhance the battery autonomy of hearing aids. The main challenge consists in optimizing the amount of energy extracted while working with soft human tissues and the shape of the earcanal. This paper proposes a new harvester concept: a liquid-filled earplug which transfers energy outside the earcanal to a generator. It is composed of a hydraulic amplifier, two hydraulic cylinders that actuate a bistable resonator to raise the source frequency driving an amplified piezoelectric transducer to generate electricity. The cycling of the resonator is achieved using two innovative flexible hydraulic valves based on the buckling of flexible tubes.\\
A multiphysics-coupled model is established to determine the system operation requirements and to evaluate its theoretical performance. This model exhibits a theoretical energy conversion efficiency of 85\%. The electromechanical performance of the resonator coupled to the piezoelectric transducer as well as the hydraulic behavior of the valves are experimentally investigated. The global model is updated with the experimental data to ameliorate its predictability to further optimize the design. The energy losses are identified, which will enable the enhance the entire design proposed and improve the experimental energy conversion efficiency to 26\%.
\end{abstract}

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress in the energy consumption of wearable mobile devices. Energy harvesting methods have been studied to complement and enhance the power supply and autonomy of batteries. In-ear devices such as hearing aids and cochlear implants are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature reaches up to $17$J for a 10 hour period of daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aid and cochlear implant users prefer disposable batteries for their length of autonomy, but would prefer rechargeable solutions if these were able to last the entire day \cite{PracticesAudiology2016}. Indeed, the long-term of rechargeable solutions makes them more economical and ecological compared to disposable cells. Such considerations have spurred the development of energy-harvesting systems.

The common sources of renewable energy are solar or wind but their use is unfeasible for the scale of this application. Another avenue comes from the fact that the human body generates a considerable amount of energy from various types of movements or conditions. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, thermal energy from body heat \cite{Kim2014}, and strain energy from skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the earcanal's mechanical deformation. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements change the earcanal's geometry during mastication. Carioli \emph{et al.} then used a customized earplug molding technique with a 3D scanner to characterize the global bending movement and local compression area in the earcanal during mastication \cite{Carioli2016}. These two mechanical deformations could potentially be exploited as an energy source.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug was molded to the subject's earcanal to maximize surface contact. The aim of the harvester was to exploit both the flexion and compression energies. The experimental prototype was able to extract $44$\,µJ from a mastication cycle. The soft materials facilitated the bio-integration of the harvester, but the PVDF's low electromechanicalcoupling coefficient may have been the cause of poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy was extracted using a liquid-filled earplug whose internal volume varied as a result of the radial compression of the earcanal during jaw movements. The earplug was equipped with a magnet in a water column surrounded by a coil. When the volume varied, the magnet would shift, generating energy. The generated energy experimentally amounted to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57\,Hz) and a small-scale application \cite{Kulah2008,Priya2017}. However, in this proposed model, the hydraulic transmission will enable the use of more complex and voluminous technological solutions since the energy will be made available outside, in other words, it will not be constricted by the earcanal's volume.

@@

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but energy can be transmitted outside the earcanal which could benefit from larger effective space. The latter is essential to be able to implement an impedance adaptation stage and more effective transducing methods that could not otherwise fit directly inside the earcanal's limited volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is proportionally greater as the effective volume of the transducing material is greater \cite{Marin2011}. Thus, the present work is based on the exploitation of the earcanal deformation using a liquid-filled earplug. As an energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the earcanal wall snugly. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Earcanal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behaviour has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12$\,kPa and $(\Delta V_{ear})_{max}=60$\,mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$\,mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$\,mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\,\% of the $17$\,J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.9\text{\,mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. In fact, the power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy sources decreases \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}.

Piezoelectric ceramics appear like good candidates for mechanical energy transduction at small scales, but their high stiffness can limit their use on soft materials as human tissues. Mechanical amplifiers like flextensional structures can nevertheless be used to adapt the stress/displacement levels between the energy source and the transducing stiff material \cite{Abdelnaby2016}. Also, the best energy conversion efficiency for piezoelectric ceramics is achieved for at the transducer resonance. As the mean mastication frequency is estimated to \mbox{$1.57$Hz}, the harvesting system could benefit from a frequency-up conversion stage \cite{Ashraf2011,Peng2021} to increase the frequency of the energy source and so the energy conversion efficiency of the transducer. A majority of the frequency-up systems with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. Stoppers however dissipate energy. Alternative solutions use mechanically bistable resonators (BR) \cite{Vocca2012}. Recent works led to the development and optimization of bistable structures integrating stacked piezoelectric ceramics inside flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling. The electromechanical converter, composed of a BR with piezoelectric ceramics in a flextensional structure, theoretically offers higher energy conversion efficiency compared to low frequency solutions and/or flexible electromechanical transducers \cite{Abdelnaby2016,Peng2021}.

The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It can store the energy in the frequency-up converter by actuating it at the frequency of human mastication. The amount of extractable energy from the hydraulic source will then depend on the mechanical stage behavior. Before being converted in electricity, the energy is in fact stored in a mechanical potential reservoir. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
	\Delta p_{ear}\ S_{hc} = K_{s}\ x_{m}
	\label{eq:hydraulic-mechanic_interface}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as follows :
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\,\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

Non-linear elastic potential reservoirs can also be considered. This work presents the energy extraction capability of a BR governed by a Duffing equation. It admits 2 stable equilibrium positions \mbox{$x_m=\pm x_0$}, an unstable one in \mbox{$x_m=0$} and its stiffness depends on its structural buckling level $\epsilon$ defined later in equation \ref{eq:epsilon_def}. The maximum theoretical energy in the BR for a quasi-static actuation is equal to the height of the potential energy barrier $E_{pb}$ when the mass position varies from \mbox{$x_m=\pm x_0$} to \mbox{$x_m=0$} (eq. \ref{eq:Ep_bar}). It will be a fraction $n$ of the upper limit of the available hydraulic energy (eq. \ref{eq:max_hydraulic_energy}) where $n$ has to be maximized. 
\begin{equation}
	E_{pb} = \dfrac{K x_0^2\epsilon^2}{3} = n(p_c \cdot (\Delta V_{ear})_{max})
	\label{eq:Ep_bar}
\end{equation}
We see later that it is inherently capable of exploiting $65$\,\% from the hydraulic energy source, which is $15$\,\% better than the best theoretical linear option. Section \ref{sec:SYSTEM MODELING} discusses the details of the modeling and design of the harvester that implements such a non-linear BR as it appeared to be a better solution than the linear approach. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the earcanal geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The figure \ref{fig:system_presentation} presents a schematic view of the energy harvesting system. It uses a liquid filled earplug to transmit the energy outside the earcanal. To maximize the efficiency, the system integrates a frequency-up conversion stage obtained by a BR. The latter is associated with an amplified piezoelectric generator (APG) \cite{CEDRATTECHNOLOGIES2022} made of a flextensional elastic APX4 steel \cite{AUBERT&DUVAL2022} structure containing stacked lead zirconate titanate (PZT) ceramics (fig. \ref{fig:APG}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.5\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={13cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.21\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The electromechanical converter is actuated, by two hydraulic cylinders (HC), with the hydraulic energy transmitted from the liquid filled earplug, through a hydraulic amplifier (HA) and two hydraulic valves (HV).

During a mastication cycle, the TMJ compresses the earcanal wall and expels the fluid out of the earplug. The HVs led the fluid alternatively to each HC in order to actuate the BR mass back and forth. The HA adapts both the cylinders stroke to the available earplug swept volume and the actuation force of the BR to the comfort pressure of the earcanal (fig. \ref{fig:system_presentation}).

The BR mass has two stable equilibrium positions, \mbox{$x_m = x_0$ }and \mbox{$-x_0$}. The harvester operates on two major phases. The first consists on the BR mass actuation, by a HC, from one stable equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass reaches and oscillates around the opposite stable equilibrium position until it stops, while a portion of the vibration energy is converted into electricity by the GPA. The mass will move backward when the next mastication occurs and complete the harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 11.5cm},clip, width=0.8\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Details of the contact between a HV and the BR mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{graphicx}
\usepackage{siunitx}

@@

\begin{abstract}
Scavenging energy from the earcanal's dynamic motion during jaw movements may be a practical way to enhance the battery autonomy of hearing aids. The main challenge consists in optimizing the amount of energy extracted while working with soft human tissues and the earcanal restricted volume. This paper proposes a new harvester concept: a liquid-filled earplug which transfers energy outside the earcanal to a generator. The latter is composed of a hydraulic amplifier, two hydraulic cylinders that actuate a bistable resonator to raise the source frequency while driving an amplified piezoelectric transducer to generate electricity. The cycling of the resonator is achieved using two innovative flexible hydraulic valves based on the buckling of flexible tubes.\\
A multiphysics-coupled model is established to determine the system operation requirements and to evaluate its theoretical performances. This model exhibits a theoretical energy conversion efficiency of 85\%. The electromechanical performance of the resonator coupled to the piezoelectric transducer, as well as the hydraulic behavior of the valves are experimentally investigated. The global model was updated using the experimental data to improve its predictability toward further optimization of the design. Moreover, the energy losses are identified to enhance the entire proposed design and improve the experimental energy conversion efficiency to 26\%.
\end{abstract}

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress in the energy consumption of wearable mobile devices. Energy harvesting methods have been studied to complement and enhance the power supply and autonomy of batteries. In-ear devices such as hearing aids and cochlear implants are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature reaches up to $17$\,J for a 10-hour period of daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select disposable batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aid and cochlear implant users prefer disposable batteries for their length of autonomy, but would prefer rechargeable solutions if these were able to last the entire day \cite{PracticesAudiology2016}. Indeed, the long-term lifespan of rechargeable solutions makes them more economical and ecological compared to disposable cells. Such considerations have spurred the development of energy-harvesting systems.

The common sources of renewable energy are solar or wind but their use is unfeasible for the scale of this application. Another avenue comes from the fact that the human body generates a considerable amount of energy from various types of movements or conditions. Any of these, the electrochemical energy from the inner ear \cite{Mercier2012}, kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, thermal energy from body heat \cite{Kim2014}, and strain energy from skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the earcanal's mechanical deformation. Delnavaz \emph{et al.} first studied in 2012 the power capability of the earcanal's geometry variation resulting from the jaw movements \cite{Delnavaz2012}. Carioli \emph{et al.} then used a customized earplug molding technique with a 3D scanner to characterize the global bending movement and local compression area in the earcanal during mastication \cite{Carioli2016}. These two mechanical deformations could potentially be exploited as an energy source.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug was molded to the subject's earcanal to maximize surface contact. The aim of the harvester was to exploit both the flexion and compression energies. The experimental prototype was able to extract $44$\,µJ from a mastication cycle. The soft materials facilitated the bio-integration of the harvester, but the PVDF's low electromechanical coupling coefficient may have been the cause of poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy was extracted using a liquid-filled earplug whose internal volume varied as a result of the radial compression of the earcanal during jaw movements. The earplug was equipped with a magnet in a water column surrounded by a coil. When the volume varied, the magnet would shift, generating energy by electromagnetic induction. The generated energy experimentally amounted to $0.2$\,µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57\,Hz) and a small-scale application \cite{Kulah2008,Priya2017}. However, in this proposed strategy, the hydraulic transmission will enable the use of more complex and voluminous technological solutions since the energy will be made available outside the earcanal.

@@

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but if energy can be transmitted outside the earcanal the device could benefit from a larger effective space. The latter is essential to the implementation of an impedance adaptation stage and more effective transducing methods that could not otherwise fit directly inside the earcanal's limited volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is proportionally greater as the effective volume of the transducing material is greater \cite{Marin2011}. Thus, the present work is based on exploiting the earcanal's deformation using a liquid-filled earplug. As an energy source, it is akin to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. To optimize energy extraction, the earplug must fit the earcanal wall snugly. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$\,kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$\,kPa of dynamic pressure variation, with a $28$\,kPa of initial pressurization, was noted during mastication for seven subjects. Furthermore, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented in Figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Earcanal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The characteristics of the liquid filled earplug's operating behavior have not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ and $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12$\,kPa and $(\Delta V_{ear})_{max}=60$\,mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user feels discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.72$\,mJ available from one mastication. If a person masticates 2200 cycles per day, which is the average according to \cite{Goll2011}, there would be $1.6$\,J available per day, per ear. This amount of energy could theoretically enhance the supply some $19$\,\% of the $17$\,J needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.72\text{\,mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. In fact, the power density of electromagnetic transducers decreases proportionately to the effective volume of the transducing material and the frequency of the energy sources decreases \cite{Priya2017,Kulah2008}. Also, the fabrication and integration of the coil and moving magnet may be a source of difficulty as the gap between the two components must be carefully managed to prevent non-linear and unpredicted responses \cite{Caruntu2001}.

Piezoelectric ceramics appear to be good candidates for small-scale mechanical energy transduction, but their rigid structure limits their use with soft materials, such as human tissue. Mechanical amplifiers, like flextensional structures, can nevertheless be used to adapt the stress/displacement levels between the energy source and the stiff transducing material \cite{Abdelnaby2016}. Also, the best energy conversion efficiency for piezoelectric ceramics is achieved at the transducer's resonance. As the mean mastication frequency is estimated to \mbox{$1.57$\,Hz}, the harvesting system could benefit from a frequency-up conversion stage to increase the frequency of the energy source, which in turn increases the energy conversion efficiency of the transducer \cite{Ashraf2011,Peng2021}. Most of the frequency-up systems with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. However, stoppers dissipate energy. Alternative solutions use mechanically bistable resonators (BR) \cite{Vocca2012}. Recent works led to the development and optimization of bistable structures integrating stacked piezoelectric ceramics within flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling. The electromechanical converter, composed of a BR with piezoelectric ceramics in a flextensional structure, theoretically offers a higher energy conversion efficiency than do low frequency solutions and/or flexible electromechanical transducers \cite{Abdelnaby2016,Peng2021}.

The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It can store the energy in the frequency-up converter by actuating it at the frequency of human mastication. The amount of extractable energy from the hydraulic source will then depend on the harvester mechanical behavior. Before being converted into electricity, the energy is in fact stored in a mechanical potential reservoir. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
	\Delta p_{ear}\ S_{hc} = K_{s}\ x_{m}
	\label{eq:hydraulic-mechanic_interface}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as follows:
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\,\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

Non-linear elastic potential reservoirs can also be considered. This work presents the energy extraction capability of a BR governed by a Duffing equation. It admits 2 stable equilibrium positions \mbox{$x_m=\pm x_0$}, an unstable one in \mbox{$x_m=0$} and its stiffness depends on its structural buckling level $\epsilon$ defined later in equation \ref{eq:epsilon_def}. The maximum theoretical energy in the BR for a quasi-static actuation is equal to the height of the potential energy barrier $E_{pb}$ when the mass position varies from \mbox{$x_m=\pm x_0$} to \mbox{$x_m=0$} (Eq. \ref{eq:Ep_bar}). It will be a fraction $n$ of the upper limit of the available hydraulic energy (Eq. \ref{eq:max_hydraulic_energy}), where $n$ has to be maximized. 
\begin{equation}
	E_{pb} = \dfrac{K x_0^2\epsilon^2}{3} = n(p_c \cdot (\Delta V_{ear})_{max})
	\label{eq:Ep_bar}
\end{equation}
Hence, we see that it is inherently capable of exploiting $65$\,\% from the hydraulic energy source, which is $15$\,\% better than the best theoretical linear option. Section \ref{sec:SYSTEM MODELING} will discuss the modeling and design of a non-linear BR harvester, as this appears to be a better solution than the linear approach. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the earcanal's geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure \ref{fig:system_presentation} is a schematic view of the energy harvesting system. In this system, a liquid-filled earplug transmits the energy beyond the earcanal. To maximize its efficiency, the system has a frequency-up conversion stage that is obtained by a BR. The BR is connected to an amplified piezoelectric generator (APG) \cite{CEDRATTECHNOLOGIES2022} made of a flextensional elastic APX4 steel \cite{AUBERT&DUVAL2022} structure that contains stacked lead zirconate titanate (PZT) ceramics (Fig. \ref{fig:APG}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.5\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={13cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.21\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The electromechanical converter is actuated by two hydraulic cylinders (HC), the hydraulic energy being transmitted from the liquid-filled earplug trough a hydraulic amplifier (HA) and two hydraulic valves (HV).

During a mastication cycle, the TMJ compresses the earcanal wall which expels the fluid out of the earplug. The HVs lead the fluid alternatively to each HC to actuate the BR mass back and forth. The HA adapts both the cylinders' stroke to the available earplug swept volume, and the actuation force of the BR to the comfort pressure of the earcanal (Fig. \ref{fig:system_presentation}).

The BR mass has two equilibrium positions, \mbox{$x_m = x_0$ } and \mbox{$-x_0$}. The harvester operates during two major phases. The first is when the BR mass is actuated by a HC, from one equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass reaches and oscillates around the opposite equilibrium position until it stops, while a portion of the vibration energy is converted into electricity by the APG. The mass will move backward when the next mastication action occurs and this completes one harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 11.5cm},clip, width=0.8\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Details of the contact between a HV and the BR mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
